Measuring causal influence with\\ back-to-back regression: the linear case

### 2019-09-25 paper/paper.tex
modifs
a

We consider the measurement of multivariate signal $Y$, generated from a set of variables $X$, via some unknown linear apparatus $F$.
%
Not all the variables in $X$ exert a causal influence on $Y$.
%
By considering a square binary diagonal matrix of \emph{causal influences} $E$, we denote by $XE$ the causal factors of $Y$.
%
In summary, the problem can be formalized as:
%
\begin{equation}
    Y = (XE + N)F,\label{eq:model}
\end{equation}
%
where $N$ is a centered, unobserved noise variable.
%
While the triplet of variables $X$ and $N$ are independent, we allow each of them to have a general covariance matrix.
% what is a general covariance matrix?
%
In practice, we observe $n$ samples $(X, Y)$ from the model.
%
This process, along with the sizes of all variables involved, is illustrated in Figure~\ref{fig:b2b}.
%
Given the model in Equation~\eqref{eq:model}, \textbf{the goal} of Back-to-Back Regression (B2B) is to estimate the matrix of causal influences $E$.

@@

Performing two regressions over the same data sample can result in overfitting, as spurious correlations in the data absorbed by the first regression will be leveraged by the second one.
%
To avoid this issue, we split our sample $(X, Y)$ at random into two subsets $(X_1, Y_1)$ and $(X_2, Y_2)$.
%
Then, the first regression is performed using $(X_1, Y_1)$, and the second regression is performed using $(X_2, Y_2)$.
%
To compensate for the reduction in sample size caused by the split, B2B is repeated over many random splits, and the final estimate $\hat E$ of the causal influence matrix is the average over the estimates associated to each split \citep{efron1992bootstrap}.
%
To accelerate this ensembling procedure, we implemented an efficient leave-one-out cross-validation scheme as detailed in \citep{rifkin2007notes} % http://cbcl.mit.edu/publications/ps/MIT-CSAIL-TR-2007-025.pdf
, as follows:
%

\begin{equation}
\hat{Y} = (\Sigma_X G Y - diag(\Sigma_X G) Y) / diag(I - \Sigma_X G)
\end{equation}

@@

We summarize the B2B procedure in Algorithm~\ref{algorithm:b2br}.
%
The rest of this section provides a theoretical guarantee on the correctness of B2B, and an optional post-processing step to binarize the estimated causal influence matrix.

@@

We perform two experiments to evaluate B2B: one on controlled synthetic data, and a second one on a real, large-scale magneto-encephalography dataset.
%
We use scikit-learn's CCA, PLS, and RidgeCV objects to compare B2B against the standard baselines \citep{sklearn}.

>>>>

We consider the measurement of multivariate signal $Y$, generated from a set of putative causes $X$, via some unknown linear apparatus $F$.
%
Not all the variables in $X$ exert a causal influence on $Y$.
%
By considering a square binary diagonal matrix of \emph{causal influences} $E$, we denote by $XE$ the causal factors of $Y$.
%
In summary, the problem can be formalized as:
%
\begin{equation}
    Y = (XE + N)F,\label{eq:model}
\end{equation}
%
where $N$ is a centered, unobserved noise variable.
%
While the triplet of variables $X$ and $N$ are independent, we allow each of them to have any form of covariance.
% what is a general covariance matrix?
%
In practice, we observe $n$ samples $(X, Y)$ from the model.
%
This problem space, along with the sizes of all variables involved, is illustrated in Figure~\ref{fig:b2b}.
%
Given the model in Equation~\eqref{eq:model}, \textbf{the goal} of Back-to-Back Regression (B2B) is to estimate the matrix of causal influences $E$.

@@

Performing two regressions over the same data sample can result in overfitting, as spurious correlations in the data absorbed by the first regression will be leveraged by the second one.
%
To avoid this issue, we split our sample $(X, Y)$ into two splits $(X_1, Y_1)$ and $(X_2, Y_2)$.
%
Then, the first regression is performed using $(X_1, Y_1)$, and the second regression is performed using $(X_2, Y_2)$.
%
To compensate for the reduction in sample size caused by the split, B2B is repeated over many random splits, and the final estimate $\hat E$ of the causal influence matrix is the average over the estimates associated to each split \citep{efron1992bootstrap}.
%
To accelerate this ensembling procedure, we implemented an efficient leave-one-out cross-validation scheme as detailed in \citep{rifkin2007notes} % http://cbcl.mit.edu/publications/ps/MIT-CSAIL-TR-2007-025.pdf
, as follows:
%

\begin{equation}
\hat{Y} = (\Sigma_X G Y - diag(\Sigma_X G) Y) \;/\; diag(I - \Sigma_X G) \qquad \text{(element-wise division)}
\end{equation}

@@

We summarize the B2B procedure in Algorithm~\ref{algorithm:b2br}.
%
The rest of this section provides a theoretical guarantee on the correctness of B2B.

@@

We perform two experiments to evaluate B2B: one on controlled synthetic data, and a second one on a real, large-scale magneto-encephalography (MEG) dataset.
%
We use scikit-learn's CCA, PLS, and RidgeCV objects to compare B2B against the standard baselines \citep{sklearn}.

### 2019-09-25 paper/synthetic.tex
a

We evaluate the performance of B2B throughout a series of experiments on
controlled synthetic data.
%
The purpose of these experiments is to evaluate the ability of B2B in terms of
prediction (inside and outside the training distribution), as well as a method
to recover causal factors.

>>>>

We evaluate the performance of B2B throughout a series of experiments on
controlled synthetic data.
%
The purpose of these experiments is to evaluate the ability of B2B in terms of
prediction of independent and identically distributed data, as well as a method
to recover causal factors.
